---
title: "README"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{knit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Earthquakes)
```

The package `Earthquakes` has been constructed to clean and plot earthquake data from NOAA.

The package is composed of three main files:

* Cleaning and formatting the dataset.
* Plotting the earthquakes on a line graph.
* Plotting the Eathquakes in a leaflet map.

## Cleaning and formatting the dataset

### A. eq_location_clean

`eq_location_clean` takes `data`, and returns that same data with its column LOCATION_NAME cleaned.

```{r, echo = FALSE}
if (require(earthquakeGeoms)){
  library(Earthquakes)
}
```

```{r, warning=FALSE}

## Function Call
data <- data.frame(LOCATION_NAME = "SPAIN:  BARCELONA")
eq_location_clean(data)
```

### B. eq_clean_data

`eq_clean_data` takes `data`  and returns a cleaned dataset with the `eq_location_clean` applied to it.


```{r}
## Data Setup
test <- data.frame(`Location Name` = c("SPAIN: BARCELONA"),
                            Year = c(2005),
                            Mo = c(5),
                            Dy = c(1),
                            Latitude = c(12.51),
                            Longitude = c(118.2),
                            Mag = c(4),
                            `Total Deaths` = c(6))
## Function Call
 eq_clean_data(test)
```


## Plotting the earthquakes on a line graph

Rendering of a line graph to plot the different earthquakes based on Richter values and deaths.


### geom_timeline

`geom_timeline` is the layer function for the GeomTime Geom. The colour indicates the number of deaths
 and size the Richter value.


```{r, fig.width=7.2, fig.height=3}
## Data Setup
test <- data.frame(`Location Name` = c("SPAIN: BARCELONA"),
                            Year = c(2005),
                            Mo = c(5),
                            Dy = c(1),
                            Latitude = c(12.51),
                            Longitude = c(118.2),
                            Mag = c(4),
                            `Total Deaths` = c(6))
## Function Call

test %>%
 eq_clean_data() %>%
 dplyr::filter(COUNTRY %in% c("TURKEY") & (lubridate::year(DATE) >= 2000 & lubridate::year(DATE) <= 2015)) %>%
 ggplot() + geom_timeline( aes(x = DATE, size = MAGNITUDE, fill = DEATHS))
 + ggplot2::labs(size = "Richter scale value", fill = "# Deaths") + theme_time()
 ```

### geom_timeline_label

`geom_timeline_label` is the layer function to call the GeomTimelineLabel Geom. This function requires
an argument  `label`. Another  argument required is `nmax`, required to know how many top labels to use in the annotation.


```{r, fig.width=7.2, fig.height=3}
## Function Call
test %>%
 eq_clean_data() %>%
 dplyr::filter(COUNTRY %in% c("TURKEY", "CALIFORNIA") & lubridate::year(DATE) > 1950) %>%
 ggplot(aes(x = DATE, y = COUNTRY, color = DEATHS, size = MAGNITUDE))
 + geom_timeline() + geom_timeline_label(aes(label =LOCATION_NAME), n_max = 3)
 + ggplot2::labs(size = "Richter scale value", col = "# Deaths") + theme_time()
```

## Plotting the Eathquakes in a leaflet map

### eq_map

`eq_map` is the function that plots the data found on the argument `data`, using the columns LATITUDE and LONGITUDE. The data will be plotted as circles in the map, once you click the circle a pop up will appear showing the information that was input in the `annot_col` argument.


```{r, fig.width=7.2, fig.height=3}

## Data Setup
test_dataframe <- data.frame(`Location Name` = c("MEXICO:  OAXACA"),                                              ),
                            Year = c(2000),
                            Mo = c(4),
                            Dy = c(4),
                            Latitude = c(17.0),
                            Longitude = c(96.3),
                            `Total Deaths`= c(100),
                            Mag  = c(7)
                            )
## Function Call
test %>%
 eq_clean_data() %>%
 dplyr::filter(COUNTRY %in% c("MEXICO") & lubridate::year(DATE) >= 2000) %>%
 eq_map(annot_col = "DATE")

```

### eq_create_label

`eq_create_label` creates a new column called 'popup_text' using location name, magnitude and total deaths.
This column is the formatted into the HTML format. If any value == NA, it is skipped. The result is a leaflet map
that contains circles representing the earthquakes, of which the size is directly proportional to the magnitude.


```{r, fig.width=7.2, fig.height=3}

## Data Setup
test_dataframe <- data.frame(`Location Name` = c("MEXICO:  OAXACA"),                                              ),
                            Year = c(2000),
                            Mo = c(4),
                            Dy = c(4),
                            Latitude = c(17.0),
                            Longitude = c(96.3),
                            `Total Deaths`= c(100),
                            Mag  = c(7)
                            )
## Function Call
test %>%
 eq_clean_data() %>%
 dplyr::filter(COUNTRY %in% c("MEXICO") & lubridate::year(DATE) >= 2000) %>%
 dplyr::mutate(popup_text = eq_create_label(.)) %>% eq_map(annot_col = "popup_text")

```
